---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## 1. Loading and preprocessing the data
```{r}
library(dplyr)
library(ggplot2)
unzip(zipfile = "C:/Users/b002882/Desktop/PRINT/NEERAJA/Reproducible Research/Week2/RepData_PeerAssessment1/repdata_data_activity.zip",exdir = "C:/Users/b002882/Desktop/PRINT/NEERAJA/Reproducible Research/Week2/RepData_PeerAssessment1")
activitydata <- read.csv("activity.csv")


```



## What is mean total number of steps taken per day?
#group by date and calculate sum of steps date wise
# 2.calculate Mean and median number of steps taken each day

```{r}
datewisesteps <- activitydata %>% select(date,steps) %>% na.omit() %>% group_by(date) %>% summarise(totalsteps=sum(steps))
mean(datewisesteps$totalsteps)
median(datewisesteps$totalsteps)

```
## 3.Histogram of the total number of steps taken each day
```{r}
hist(datewisesteps$totalsteps, xlab="Total Steps",main="Total steps taken per day ", breaks = 20)


```
## 4.What is the average daily activity pattern?

```{r}
dailyactivityinterval <- activitydata %>% select(interval,steps) %>% na.omit() %>% group_by(interval) %>% summarize(meansteps=mean(steps))
ggplot(dailyactivityinterval,aes(x=interval, y=meansteps))+geom_line()+ggtitle("Average daily activity pattern")

```

## 5.The 5-minute interval that, on average, contains the maximum number of steps

```{r}
dailyactivityinterval[which(dailyactivityinterval$meansteps == max(dailyactivityinterval$meansteps)),]
```
## 6.Imputing missing values
#6.1.Calculate and report the total number of missing values in the dataset 

```{r}
sum(is.na(activitydata$steps))

```
#6.2.strategy:fill  all of the missing values in the dataset by the mean for that 5-minute interval.
# cbind dataframe with missing steps data and meansteps calculated dataframe

```{r}
mergeActivity <- cbind(activitydata,dailyactivityinterval)

```
#6.3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
mergeActivity$steps[is.na(mergeActivity$steps)]<- mergeActivity$meansteps[is.na(mergeActivity$steps)]
```
#6.4.histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r}
imputedatewisesteps <- mergeActivity %>% select(date,steps) %>% group_by(date) %>% summarise(totalsteps=sum(steps))

hist(imputedatewisesteps$totalsteps, xlab="Total Steps",main="Total steps taken per day ", breaks = 20)

```
#6.4.1.Calculate and report the mean and median total number of steps taken per day
```{r}
mean(imputedatewisesteps$totalsteps)
median(imputedatewisesteps$totalsteps)
```
## Are there differences in activity patterns between weekdays and weekends?
```{r}

weekdetails <- subset(mergeActivity[,c("steps","date","interval")])

weekdetails <- transform(weekdetails, date=as.Date(date))

weekdetails$weekday <- weekdays(weekdetails$date)

weekdetails$week  <- ifelse(weekdetails$weekday=="Saturday" |weekdetails$weekday=="Sunday","Weekend","Weekday")

ggplot(weekdetails,aes(x=interval,y=steps,color=week)) + geom_line()+facet_grid(week~.)+xlab("Interval")+ylab("Steps")+ggtitle("Activity patterns in weekdays and weekends")

```

